---
title: "uscovid19vacc Tutorial"
output: rmarkdown::html_vignette
description: This explains how to use the uscovid19vacc package to pull data collected by the Johns Hopkins Infectious Disease Dynamics Group and Center for Civic Impact, and clean and standardize it.
vignette: >
  %\VignetteIndexEntry{uscovid19vacc Tutorial}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
```


GET VACCINATION BY AGE AND STATE, USA

```{r, message = FALSE}
library(tidyverse)
library(uscovid19vacc)
#source("R/vacc_funcs.R")
```




Set the github token for the data, since it's currently a private repo.

```{r, message = FALSE}
git_token <- "AB63TGGKM6VBDNGJHXRRP7LAYPDIM"
```



## Current Vaccination Coverage

```{r}

daily_state_vacc <- get_state_vacc() %>%
    filter(!is.na(dose1))

```


## Pull and Clean Data
Pull data from JHU CCI github and combine with JHU IDD data that is part of the package.

```{r}

# Get clean, but not age group standardized data
vacc_data <- get_clean_us_agevacc(git_token = git_token)

# Make it standardized to either 5yr or 10 year age groups
vacc_data_5yr <- get_standardized_us_agevacc(vacc_data = vacc_data,
                                        age_groups = "5yr", # "10yr"
                                        git_token = git_token)

# Make it standardized to either 5yr or 10 year age groups
vacc_data_10yr <- get_standardized_us_agevacc(vacc_data = vacc_data,
                                        age_groups = "10yr", 
                                        git_token = git_token) 


```



## Plot a sample of states to check

```{r, fig.width=8, fig.height=8}
usps_samp <- sample(unique(vacc_data_5yr$USPS), 12, replace = FALSE)
vacc_data_5yr %>% filter(USPS %in% usps_samp) %>%
              arrange(prop_vacc_age) %>%
              mutate(USPS = fct_reorder(USPS, (prop_vacc_age))) %>%
              ggplot(aes(x=date, y=prop_vacc_age, color=age_group, group=age_group)) +
              geom_line() +
              geom_point(aes(shape=source, fill=source)) +
              scale_shape_manual("Source", values = c(21, 23)) +
              scale_fill_manual("Source", values=c("orange","lightblue4")) +
              scale_color_discrete("Age group, yrs") +
              xlab("State") + ylab("proportion vaccinated") +
              ggtitle("Proportion Vaccinated, By State and Age") +
              theme_bw() +
              theme(axis.text.x = element_text(angle=90)) +
              facet_wrap(~USPS)
```

```{r fig.width=8, fig.height=8}
vacc_data_10yr %>% filter(USPS %in% usps_samp) %>%
              arrange(prop_vacc_age) %>%
              mutate(USPS = fct_reorder(USPS, (prop_vacc_age))) %>%
              ggplot(aes(x=date, y=prop_vacc_age, color=age_group, group=age_group)) +
              geom_line() +
              geom_point(aes(shape=source, fill=source)) +
              scale_shape_manual("Source", values = c(21, 23)) +
              scale_fill_manual("Source", values=c("orange","lightblue4")) +
              scale_color_discrete("Age group, yrs") +
              xlab("State") + ylab("proportion vaccinated") +
              ggtitle("Proportion Vaccinated, By State and Age") +
              theme_bw() +
              theme(axis.text.x = element_text(angle=90)) +
              facet_wrap(~USPS)
```



## Generate plots of all 

### 5-year age groups
```{r}
dir.create("figures", recursive = TRUE, showWarnings = FALSE)
states_ <- sort(unique(vacc_data_5yr$USPS))
pages_ <- ceiling(length(states_)/9)
state_groups <- sort(rep(1:pages_, 9))[1:length(states_)]

pdf(paste0("figures/prop_vacc_by_age_", Sys.Date(), ".pdf"), height=11, width = 8.5)
for (i in 1:pages_){
    state_tmp <- states_[state_groups==i]
    print(vacc_data_5yr %>% filter(USPS %in% state_tmp) %>%
              arrange(prop_vacc_age) %>%
              mutate(USPS = fct_reorder(USPS, (prop_vacc_age))) %>%
              ggplot(aes(x=date, y=prop_vacc_age, color=age_group, group=age_group)) +
              geom_line() +
              geom_point(aes(shape=source, fill=source)) +
              scale_shape_manual("Source", values = c(21, 23)) +
              scale_fill_manual("Source", values=c("orange","lightblue4")) +
              scale_color_discrete("Age group, yrs") +
              xlab("State") + ylab("proportion vaccinated") +
              ggtitle("Proportion Vaccinated, By State and Age") +
              theme_bw() +
              theme(axis.text.x = element_text(angle=90)) +
              facet_wrap(~USPS))
}
dev.off()

```


### 10-year age groups

```{r}
states_ <- sort(unique(vacc_data_10yr$USPS))
pages_ <- ceiling(length(states_)/9)
state_groups <- sort(rep(1:pages_, 9))[1:length(states_)]

pdf(paste0("figures/prop_vacc_by_age10yr_", Sys.Date(), ".pdf"), height=11, width = 8.5)
for (i in 1:pages_){
    state_tmp <- states_[state_groups==i]
    print(vacc_data_10yr %>% filter(USPS %in% state_tmp) %>%
              arrange(prop_vacc_age) %>%
              mutate(USPS = fct_reorder(USPS, (prop_vacc_age))) %>%
              ggplot(aes(x=date, y=prop_vacc_age, color=age_group, group=age_group)) +
              geom_line() +
              geom_point(aes(shape=source, fill=source)) +
              scale_shape_manual("Source", values = c(21, 23)) +
              scale_fill_manual("Source", values=c("orange","lightblue4")) +
              scale_color_discrete("Age group, yrs") +
              xlab("State") + ylab("proportion vaccinated") +
              ggtitle("Proportion Vaccinated, By State and Age") +
              theme_bw() +
              theme(axis.text.x = element_text(angle=90)) +
              facet_wrap(~USPS))
}
dev.off()

```


## Clean dates that are problematic

```{r}

vacc_data_5yr_cl <- drop_dates_vaccdat(vacc_data_Xyr = vacc_data_5yr)
vacc_data_10yr_cl <- drop_dates_vaccdat(vacc_data_Xyr = vacc_data_10yr)

```

## Plot a sample of states

```{r fig.width=8, fig.height=8}
vacc_data_10yr_cl %>% filter(USPS %in% usps_samp) %>%
              arrange(prop_vacc_age) %>%
              mutate(USPS = fct_reorder(USPS, (prop_vacc_age))) %>%
              ggplot(aes(x=date, y=prop_vacc_age, color=age_group, group=age_group)) +
              geom_line() +
              geom_point(aes(shape=source, fill=source)) +
              scale_shape_manual("Source", values = c(21, 23)) +
              scale_fill_manual("Source", values=c("orange","lightblue4")) +
              scale_color_discrete("Age group, yrs") +
              xlab("State") + ylab("proportion vaccinated") +
              ggtitle("Proportion Vaccinated, By State and Age") +
              theme_bw() +
              theme(axis.text.x = element_text(angle=90)) +
              facet_wrap(~USPS)
```





## Generate plots of cleaned data

### 5-year age groups

```{r}
states_ <- sort(unique(vacc_data_5yr_cl$USPS))
pages_ <- ceiling(length(states_)/9)
state_groups <- sort(rep(1:pages_, 9))[1:length(states_)]

pdf(paste0("figures/prop_vacc_by_age5yr_", Sys.Date(), "_clean.pdf"), height=11, width = 8.5)
for (i in 1:pages_){
    state_tmp <- states_[state_groups==i]
    print(vacc_data_5yr_cl %>% filter(USPS %in% state_tmp) %>%
              arrange(prop_vacc_age) %>%
              mutate(USPS = fct_reorder(USPS, (prop_vacc_age))) %>%
              ggplot(aes(x=date, y=prop_vacc_age, color=age_group, group=age_group)) +
              geom_line() +
              geom_point(aes(shape=source, fill=source)) +
              scale_shape_manual("Source", values = c(21, 23)) +
              scale_fill_manual("Source", values=c("orange","lightblue4")) +
              scale_color_discrete("Age group, yrs") +
              xlab("State") + ylab("proportion vaccinated") +
              ggtitle("Proportion Vaccinated, By State and Age") +
              theme_bw() +
              theme(axis.text.x = element_text(angle=90)) +
              facet_wrap(~USPS))
}
dev.off()

```


### 10-year age groups

```{r}
dir.create("figures", recursive = TRUE, showWarnings = FALSE)
states_ <- sort(unique(vacc_data_10yr_cl$USPS))
pages_ <- ceiling(length(states_)/9)
state_groups <- sort(rep(1:pages_, 9))[1:length(states_)]

pdf(paste0("figures/prop_vacc_by_age10yr_", Sys.Date(), "_clean.pdf"), height=11, width = 8.5)
for (i in 1:pages_){
    state_tmp <- states_[state_groups==i]
    print(vacc_data_10yr_cl %>% filter(USPS %in% state_tmp) %>%
              arrange(prop_vacc_age) %>%
              mutate(USPS = fct_reorder(USPS, (prop_vacc_age))) %>%
              ggplot(aes(x=date, y=prop_vacc_age, color=age_group, group=age_group)) +
              geom_line() +
              geom_point(aes(shape=source, fill=source)) +
              scale_shape_manual("Source", values = c(21, 23)) +
              scale_fill_manual("Source", values=c("orange","lightblue4")) +
              scale_color_discrete("Age group, yrs") +
              xlab("State") + ylab("proportion vaccinated") +
              ggtitle("Proportion Vaccinated, By State and Age") +
              theme_bw() +
              theme(axis.text.x = element_text(angle=90)) +
              facet_wrap(~USPS))
}
dev.off()

```


Save the data

```{r}
dir.create("output")
readr::write_csv(vacc_data_5yr_cl, "output/vacc_data_5yr_cl.csv")
readr::write_csv(vacc_data_10yr_cl, "output/vacc_data_10yr_cl.csv")

```





